---
title: "IXIS Data challenge"
author: "Kevin Duan"
date: "7/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(openxlsx)
library(ggplot2)
library(lubridate)
library(reshape2)
```

## Data Acquisition

```{r}
#acquire the data from csv
sessionCounts <- read.csv("C:/Users/Kevin Duan/Downloads/DataAnalyst_Ecom_data_sessionCounts.csv")

cart <- read.csv("C:/Users/Kevin Duan/Downloads/DataAnalyst_Ecom_data_addsToCart.csv")
```

## Data Exploration


```{r exploration}
head(sessionCounts)
head(cart)

#change the date variable to date datatype
sessionCounts$dim_date <- as.Date(sessionCounts$dim_date, format = "%m/%d/%y")
```
The datasets contain 6 columns. The data set includes:

* `dim_browser`(categorical) which identifies the browser used
* `dim_deviceCategory` (categorical) which identifies device used
* `dim_date` (date) which identifies date site was accessed or date of transaction(?)
* `sessions`(int) which shows how many sessions the user was on the site 
* `transactions`(int) which signifies transactions made
* `QTY` (int) which shows the amount purchased
* `addsToCart` (int) shows the quantity items were added to cart on the site. 
* `dim_year` and `dim_month` are self explanatory

From the dataset, explore to determine distribution and look for any missing values/outliers


```{r exploration2}
#explore the distribution of the quantitative features
summary(sessionCounts[,c(3:6)])

#explore the distributions of the categorical features
table(sessionCounts$dim_browser, sessionCounts$dim_deviceCategory)

#the table version is difficult to see because of so many different browsers, we plot the data instead to visualize it.
ggplot(sessionCounts, aes(x= `dim_browser`)) + 
  geom_bar() +
  xlab("Browser types") +
  theme(
    panel.background = element_blank(),
    axis.text.x = element_text(angle = 90,hjust =0.95, vjust = 0.5)
  ) +
  coord_flip()

prop.table(table(sessionCounts$dim_deviceCategory))
length(unique(sessionCounts$dim_browser))
unique(sessionCounts$dim_browser)
sessionCounts %>% filter(dim_browser == c( "DESKTOP"))


#We also explore the cart dataset as well
summary(cart[3])

```

From the exploration above, we found that

* the data ranges from 7/1/12 to 06/30/13 
* the `sessions` data is right skewed and has a large maximum which largely affects the mean. The `transactions` and `QTY` data follow a similar trend.
* There are many different browsers used (**57**): some repeating, some which aren't browsers, and some which aren't labelled.
* the visitation of the sights are mobile (**39%**), desktop (**35%**), and tablet(**26%**) in order of most sessions to the least.
* The site got the most cart additions on 8/2012, and the least on the last month 06/2013.

Based on the exploration done thus far, we could assume that the website has difficulty completing transactions, other than some large outliers.
We may want to pivot the device feature into separate columns. We also may want to combine/delete some of the browser types as they are duplicates or not usable.

## Data transformation and Cleaning

```{r transformation}
(sheet1 <- sessionCounts %>%
   #remove the (notset) and error browser types. After looking on google analytics (GA) it said that these could be tag errors. 
  filter(!(dim_browser %in% c('(not set)','error'))) %>%
  #subset the data to the relevant columns for the client
  select(dim_deviceCategory, dim_date, sessions, transactions, QTY) %>%
  #aggregate the data by months and devices used
  group_by(month = lubridate::floor_date(dim_date, 'month'), device = dim_deviceCategory) %>%
   summarise(sessions = sum(sessions), QTY = sum(QTY), transactions = sum(transactions), ECR = sum(transactions)/sum(sessions))
)

(newCounts <- sessionCounts %>%
    #use only the last 2 month's data
    #also remove the (notset) and error browser types
    filter(dim_date >= "2013-05-01" & dim_date <= "2013-06-30" & !(dim_browser %in% c('(not set)','error'))) %>%
    
  #group the data set by months to be joined with the carts data
  group_by(date = lubridate::floor_date(dim_date, 'month')) %>%
    
    #sum up all of the metrics by month
    summarise(sessions = sum(sessions), QTY = sum(QTY), transactions = sum(transactions)) %>%
    
    #arrange the order of new tibble by month, and then sessions
    arrange(date, desc(sessions))%>%
    #rename the month values so they can be joined with the carts dataset
    mutate(date=as.character(format(date, "%Y-%m")))
)
    
#join the columns to format them the same as the sheet 2 table
(joinCart <- cart %>%
    unite(date, c(dim_year, dim_month), sep = "-0" ) %>%
    filter(date %in% c("2013-05","2013-06"))
)  

#create a relative difference function
rdiff <- function(x){
  reldiff<- diff(x)/x[-length(x)]
}

#join the two tables
(sheet2 <- merge(newCounts, joinCart, all.x = TRUE) %>%
  #create rows that measure the difference between the two months
  bind_rows(summarise(.,
                      across(where(is.numeric),diff),
                      across(where(is.character), ~"absdiff")), 
            summarise(.,
                      across(where(is.numeric), rdiff),
                      across(where(is.character), ~"reldiff")))
)
```
```{r}
#export the two tables as excel worksheets
ixisworksheet <- list('sheet1' = sheet1, 'sheet2' = sheet2)
write.xlsx(ixisworksheet, file = "C:/Users/Kevin Duan/Documents/GitHub/IXISDataChallenge/IXISDataChallenge.xlsx")
```


```{r echo=FALSE}
sheet1 %>% ggplot(aes(x = month, y=(ECR), color = device)) + 
  geom_line(size = 1.2) + 
  labs(x = "Months", y = "ECR", color = "Devices",
       caption = "ECR = Transactions/Sessions", title = " Website Trends over 12 Months")+
theme(panel.background = element_blank(),
      panel.grid.major.x = element_line(colour = "lightgray"),
      ) +
  scale_x_date(breaks = (function(x) seq.Date(from = as.Date("2012-07-01"), to=as.Date("2013-06-30") , by="2 months")))

#website trends
sheet1 %>% filter(device == "desktop") %>%
  ggplot()+
  geom_line(aes(x= month, y= sessions)) +
theme(panel.background = element_blank(),
      panel.grid.major= element_line(colour = "lightgray")
      )+
  labs(title = "Web Traffic for Desktop Users Over 12 Months", y = "Sessions", x= "Months")

sheet1 %>% filter(device == "desktop") %>%
  ggplot()+
  geom_line(aes(x= month, y= transactions)) +
theme(panel.background = element_blank(),
      panel.grid.major= element_line(colour = "lightgray")
      )+
  labs(title = "Web Transactions for Desktop Users Over 12 Months", y = "Transactions", x= "Months")

```
